
```{r}
library(tidyverse)
library(censusapi)
```

## Look at available end points


```{r}
census_apis <- listCensusApis() %>% view()
```


## Stratify by sex for those under 5 at census tract level

```{r}
listCensusMetadata(
    name = "acs/acs5/spt",
    vintage = 2015,
    type = "variables"
) %>% view()
```

B01001_003 = Estimate!!Total:!!Male:!!Under 5 years
B01001_027 = Estimate!!Total:!!Female:!!Under 5 years

## Stratify by sex for those under 5 years old at county level

```{r}
listCensusMetadata(
    name = "acs/acs5/subject",
    vintage = 2019,
    type = "variables"
) %>% view()
```

S0101_C01_002E = Estimate!!Total!!Total population!!AGE!!Under 5 years
S0101_C05_002E = Estimate!!Female!!Total population!!AGE!!Under 5 years
S0101_C03_002E = Estimate!!Male!!Total population!!AGE!!Under 5 years

```{r}
age_sex_population_estimates <- 
    censusapi::getCensus(
        name = "acs/acs5/subject",
        vintage = 2019,
        region = "county:031",
        regionin = "state:17",
        vars = c("S0101_C01_002E", "S0101_C05_002E", "S0101_C03_002E")
    )

age_sex_population_estimates %>%
    rename(
        female = S0101_C05_002E,
        male = S0101_C03_002E
    )

view(age_sex_population_estimates)
```

## Stratify by race/ethnicity for those under 5 years old at county level


```{r}
listCensusMetadata(
    name = "acs/acs1/spp",
    vintage = 2019,
    type = "variables"
) %>% view()
```

S0201_006E = Estimate!!SEX AND AGE!!Total population
S0201_009E = Estimate!!SEX AND AGE!!Total population!!Under 5 years
POPGROUP = Race/Ethnic Group

```{r}
age_race_ethn_pop_estimates <- 
    censusapi::getCensus(
        name = "acs/acs1/spp",
        vintage = 2019,
        region = "county:031",
        regionin = "state:17",
        vars = c("S0201_006E", "S0201_009E", "POPGROUP", "POPGROUP_LABEL")
    )

age_race_ethn_pop_estimates_mod_1 <- 
    age_race_ethn_pop_estimates %>%
    rename(
        total_of_race_eth = S0201_006E,
        percent_of_race_eth_under_5 = S0201_009E
    ) %>%
    mutate(
        count_of_race_eth_under_5 = percent_of_race_eth_under_5 / 100 * total_of_race_eth
    )

total_pop_under_5 <- 
    age_race_ethn_pop_estimates_mod_1$count_of_race_eth_under_5 %>%
    pluck(1)

age_race_ethn_pop_estimates_mod_1 %>%
    mutate(
        percent_of_under_5_w_race_eth = round(count_of_race_eth_under_5 / total_pop_under_5 * 100, 2)
    ) %>%
    view()
```

## County vs City

For ACS 2019 5-year Estimate

```{r}
city_under_5 <- 171323
county_under_5 <- 310231
county_suburbs_under_5 <- county_under_5 - city_under_5
city_under_5 / county_under_5
county_suburbs_under_5 / county_under_5
```

