---
title: "Infant Asphyxiation Mortality Modeling"
author: "Daniel P. Riggins, MD"
date: "11/1/2021"
output: html_document
---

This modeling analysis will seek to predict infant asphyxiation deaths in Chicagoland census tracts using a variety of census-level predictor variables, many of which are derived from the CDC's [Social Vulnerability Index (SVI)](https://seandavi.github.io/sars2pack/reference/cdc_social_vulnerability_index.html).

# Set Up

## Load needed libraries:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(DataExplorer)
library(corrr)
library(vcd)
library(tidycensus)
library(sf)
tidymodels_prefer()
```

## Set pseudo-random seed for reproducibility:

```{r}
set.seed(324)
```

## Import the data:

```{r}
raw <- readxl::read_xlsx("data/finaldataforanalysis3_110221.xlsx")

glimpse(raw)
```

# Get census tract populations and spatial features

```{r}
enriched <-
    get_acs(
        geography = "tract",
        variables = "B01003_001",
        state = "IL",
        geometry = TRUE
    ) %>%
    filter(GEOID %in% raw$FIPS) %>%
    mutate(FIPS = as.numeric(GEOID)) %>%
    select(
        FIPS,
        population = estimate
    ) %>%
    full_join(raw)

glimpse(enriched)
```

# First Pass Exploration

## Split data into training and test sets:

```{r}
split <- initial_split(enriched, prop = 0.75)
training <- training(split)
testing <- testing(split)
```


## Summarize the primary outcome

```{r}
table(training$Count_Asphyxia)
```
## Visualize the primary outcome:

```{r}
plot_histogram(training$Count_Asphyxia)
```

The distribution of outcome data seems consistent with a count outcome approximated by theoretical distributions like the Poisson or negative binomial.

## Visualize raw predictors

```{r}
plot_histogram(select(training, 2:32))
```

We can anticipate that many of these variables will be highly correlated with each other. Let's look closer.

    )

summary(tbl$count_Asphyxia_cat)
```

```{r}
plot_intro(tbl)
```


