### Ingestion

At time of primary analysis, DPR combined a data file generated by collaborator HZ and additional data pulled via the R package {[tidycensus](https://walker-data.com/tidycensus/index.html)}.

DPR imported HZ's excel file into the Targets pipeline using the function `readxl::read_xls()`. He pulled from the tidycensus API using his function `get_suid_from_tidycensus()`. This function asks for variables from the American Community Survey's 5-year estimate for 2015-2019. The variables describe count of total population, count of population under age 5 years, average number of people per household, and geospatial shape of census tracts in Cook County, IL. The function also does some reshaping of data as noted in the code comments below:

```{r echo=FALSE}
print(source("R/get_suid_from_tidycensus_raw.R")$value)
```

DPR combined the two data sources with his function `assemble_suid()`, which performs a left join on FIPS ID for observations from HZ's excel data, then does some wrangling around naming conventions, adds a boolean variable for whether SUID is present in a given tract, and rounds estimates to the first decimal point for consistency:

```{r echo=FALSE}
print(source("R/assemble_suid.R")$value)
```